\documentclass[a4paper,12pt]{article}

% Use the Classic Thesis style
%
% \usepackage[nochapters]{classicthesis}

\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}

% Use text fonts
%
\usepackage{lmodern}  % the Latin Modern fonts which is an enhanced version of the Computer Modern fonts
% \usepackage{tgschola}  % the Gyre Schola fonts which is an enhanced version of the New Centrury Schoolbook fonts

% Use math fonts
%
% \usepackage{concmath}  % the Concrete Math fonts
% \usepackage{euler}  % the Euler Math fonts
% \usepackage{kerkis}  % the Kerkis fonts

% Some font packages that provide math font may override the text font, the
% following three commands reset it to the default.
%
% For the Latin Modern fonts
%
% \renewcommand{\rmdefault}{lmr}
% \renewcommand{\sfdefault}{lmss}
% \renewcommand{\ttdefault}{lmtt}
%
% For the Gyre Schola fonts
%
% \renewcommand{\rmdefault}{qcs}

\usepackage{amsmath}

% Use the MH bundle
%
% \usepackage{mhsetup}
% \usepackage{mathtools}
% \usepackage{mathstyle}
% \usepackage{breqn}
% \usepackage{empheq}
% \usepackage{flexisym}

% Use AMS-style theorem
%
\usepackage{amsthm}
% \theoremstyle{definition}
% \newtheorem{theorem}{Theorem}

% Use MATH formulas in PARagraph mode Typesetting Inference Rules
% 
\usepackage{mathpartir}

% \usepackage{tikz}
% \usepackage{tikz-qtree}
% \usepackage{bigfoot}

% New commands
%
\newcommand{\term}[1]{\textsf{#1}}
\newcommand{\appl}[2]{#1\inparens{#2}}
\newcommand{\eval}[2]{#1 \Longrightarrow #2}
\newcommand{\redc}[2]{#1 \longrightarrow #2}

\include{prelude}

% Front elements
%
\title{
 Programming Languages and Types \\~\\
 \textbf{Homework 11}
}
\author{
 Yi Dai
}

\begin{document}

\maketitle


\section{$\lambda$-Calculus vs. $\lambda$-Based Language}

\begin{enumerate}
 \item Summarize the differences between $\lambda$-calculus and functional programming
	languages based on it.
\end{enumerate}

\section{Nameless $\lambda$-Calculus}

\begin{enumerate}
 \item Implement a function \verb|removeNames| that takes a naming context and a named
  $\lambda$-term and returns the	corresponding nameless term.
 \item Implement a function \verb|regainNames| that takes a naming context and a nameless
  $\lambda$-term and returns a named term.
 \item{} [\emph{optional}] Implement the \verb|shift| and \verb|subst| function.
 \item{} [\emph{optional}] Implement the evaluation semantics for $\lambda_{bv}$ using
  de-Bruijn indices.
\end{enumerate}

\noindent
Note that your implementations for \verb|removeNames| and \verb|regainNames| should be
inverse of each other.

\section{Substitution vs. Environment}

\begin{enumerate}
 \item Implement the substitution-based evaluation semantics for $\lambda_{ao}$ using
  de-Bruijn indices.
 \item Implement the environment-based evaluation semantics for $\lambda_{bn}$ using
  de-Bruijn indices.
\end{enumerate}

\section{Computational $\lambda$-Calculus}

\begin{enumerate}
 \item Implement the type-checking algorithm for the simply-typed $\lambda$-calculus
  with booleans.
 \item{} [\emph{optional}] Add support for monads to your type-checking algorithm.
\end{enumerate}

\end{document}
